taxize - taxonomic search and retrieval in R

### 2013-05-10 inst/doc/taxize_withcode.tex
cached files, small changes to text, better graphics

\hlcomment{# let's take a peek at the data, excluding the data source ID and score}
\hlcomment{# columns}
temp[, -\hlfunctioncall{c}(1, 4)]
\end{alltt}
\begin{verbatim}
    submitted_name                 name_string                   title
1 Helianthos annus        Helianthus annuus L.       Catalogue of Life
3 Helianthos annus            Helianthus annus GBIF Taxonomic Backbone
4 Helianthos annus            Helianthus annus                     EOL
5 Helianthos annus         Helianthus annus L.                     EOL
6 Helianthos annus            Helianthus annus           uBio NameBank
2    Homo saapiens Homo sapiens Linnaeus, 1758       Catalogue of Life
\end{verbatim}
\end{kframe}
\end{knitrout}

Looks like the correct spellings are \emph{Helianthus annuus} and \emph{Homo sapiens}, cool!

Another approach is using the Taxonomic Name Resolution Service via the Taxosaurus API (\url{http://taxosaurus.org/}).

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# Lets set our list of species names}
mynames <- \hlfunctioncall{c}(\hlstring{"Helianthus annuus"}, \hlstring{"Pinus contort"}, \hlstring{"Poa anua"}, \hlstring{"Abis magnifica"}, 
    \hlstring{"Rosa california"}, \hlstring{"Festuca arundinace"}, \hlstring{"Sorbus occidentalos"}, \hlstring{"Madia sateva"})

\hlcomment{# And we'll call the API with the tnrs function}
\hlfunctioncall{tnrs}(query = mynames)[, -\hlfunctioncall{c}(5:7)]
\end{alltt}
\begin{verbatim}
data frame with 0 columns and 0 rows
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

>>>>

\hlcomment{# Let's take a peek at the data, excluding the data source ID and score}
\hlcomment{# columns}
temp[, -\hlfunctioncall{c}(1, 4)]
\end{alltt}
\end{kframe}
\end{knitrout}

It looks like the correct spellings are \emph{Helianthus annuus} and \emph{Homo sapiens}. Another approach uses the Taxonomic Name Resolution Service via the Taxosaurus API (\url{http://taxosaurus.org/}).

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# A list of species names, some of which are misspelled}
mynames <- \hlfunctioncall{c}(\hlstring{"Helianthus annuus"}, \hlstring{"Pinus contort"}, \hlstring{"Poa anua"}, \hlstring{"Abis magnifica"}, 
    \hlstring{"Rosa california"}, \hlstring{"Festuca arundinace"}, \hlstring{"Sorbus occidentalos"}, \hlstring{"Madia sateva"})

\hlcomment{# And we'll call the API with the tnrs function, and remove a few columns}
\hlfunctioncall{tnrs}(query = mynames)[, -\hlfunctioncall{c}(5:7)]
\end{alltt}
\begin{verbatim}
        submittedName        acceptedName    sourceId score
7   Helianthus annuus   Helianthus annuus iPlant_TNRS  1.00
4       Pinus contort      Pinus contorta iPlant_TNRS  0.98
5            Poa anua           Poa annua iPlant_TNRS  0.96
3      Abis magnifica     Abies magnifica iPlant_TNRS  0.96
8     Rosa california    Rosa californica iPlant_TNRS  0.99
2  Festuca arundinace Festuca arundinacea iPlant_TNRS  0.99
1 Sorbus occidentalos Sorbus occidentalis iPlant_TNRS  0.99
6        Madia sateva        Madia sativa iPlant_TNRS  0.97
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\subsection{Interactive name selection}

In the previous section we used the function \emph{get_tsn} to get a classification for two plant species.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# In this example, this search term has many matches. The function returns}
\hlcomment{# a data.frame of the matches, and asks for user input. Enter the row}
\hlcomment{# number of the name you want.}
\hlfunctioncall{get_tsn}(searchterm = \hlstring{"Elisabethae"}, searchtype = \hlstring{"sciname"})
\end{alltt}
\begin{verbatim}

Retrieving data for species ' Elisabethae '

                           combinedname    tsn
1                  Bittacus elisabethae 810495
2            Caenaugochlora elisabethae 754498
3                  Ceratina elisabethae 765455
4            Dolicheremaeus elisabethae 740429
5                  Euphione elisabethae  64893
6                 Harmothoe elisabethae 573734
7                  Hydraena elisabethae 815669
8                    Lizzia elisabethae  48875
9                   Melinna elisabethae  67765
10            Myrmicocrypta elisabethae 580527
11           Myzomela boiei elisabethae 713955
12             Notosacantha elisabethae 838430
13               Odontomyia elisabethae 627043
14             Parabelbella elisabethae 737749
15                Peltigera elisabethae 191230
16          Polymorphanisus elisabethae 600292
17        Pseudopterogorgia elisabethae  52264
18                  Rhantus elisabethae 815107
19              Scyllarides elisabethae 553004
20       Sellnickochthonius elisabethae 735137
21       Serilophus lunatus elisabethae 711636
22                Stenhelia elisabethae  87927
23              Tetramorium elisabethae 582397
24                Theridion elisabethae 868794
25             Trichoniscus elisabethae 597365
26 Trichoniscus elisabethae estoniensis 598124

More than one TSN found for species ' Elisabethae '!

          Enter rownumber of species (other inputs will return 'NA'):
\end{verbatim}
\begin{alltt}

\hlcomment{# You can pass in a long list of taxonomic names}
splist <- \hlfunctioncall{c}(\hlstring{"annona cherimola"}, \hlstring{"annona muricata"}, \hlstring{"quercus robur"}, \hlstring{"shorea robusta"}, 
    \hlstring{"pandanus patina"}, \hlstring{"oryza sativa"}, \hlstring{"durio zibethinus"})
\hlfunctioncall{get_tsn}(searchterm = splist, searchtype = \hlstring{"sciname"})
\end{alltt}
\begin{verbatim}

Retrieving data for species ' annona cherimola '

Retrieving data for species ' annona muricata '

Retrieving data for species ' quercus robur '

Retrieving data for species ' shorea robusta '

Retrieving data for species ' pandanus patina '

Retrieving data for species ' oryza sativa '

Retrieving data for species ' durio zibethinus '
[1] "506198" "18098"  "19405"  "506787" "507376" "41976"  "506099"
attr(,"class")
[1] "tsn"
\end{verbatim}
\begin{alltt}

\hlcomment{# In this example, note that no match at all returns an NA}
\hlfunctioncall{get_uid}(sciname = \hlfunctioncall{c}(\hlstring{"Chironomus riparius"}, \hlstring{"aaa vva"}))
\end{alltt}
\begin{verbatim}

Retrieving data for species ' Chironomus riparius '

Retrieving data for species ' aaa vva '
[1] "315576" NA      
attr(,"class")
[1] "uid"
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
text <- \hlstring{"9,562445749\textbackslash{}n37,66119516\textbackslash{}n72,9103195\textbackslash{}n"}
\hlfunctioncall{scan}(dec = \hlstring{","}, text = text)
\end{alltt}
\begin{verbatim}
[1]  9.562 37.661 72.910
\end{verbatim}
\end{kframe}
\end{knitrout}

### 2013-05-10 inst/doc/taxize_withcode_appA.Rnw
cached files, small changes to text, better graphics

\documentclass[letterpaper,superscriptaddress,showkeys,longbibliography,10pt]{revtex4-1}
\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\usepackage{natbib}
\usepackage{footnote}
\hypersetup
{
colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

@@

<<resolvenames, eval=TRUE>>=
# The tnrs function accepts a vector of 1 or more
splist_tnrs <- tnrs(query=splist, getpost="POST", source_ = "iPlant_TNRS")

@@

<<commonnames, eval=TRUE>>=
tsns <- get_tsn(searchterm=splist, searchtype="sciname", verbose=FALSE)
comnames <- lapply(tsns, getcommonnamesfromtsn)

@@

Another common task is getting the taxonomic tree upstream from your study taxa. We often know what family or order our taxa are in, but it we often don't know the tribes, subclasses, and superfafamilies. taxize provides many avenues to getting classifications. Two of them are accesible via a single function (\emph{classification}): the Integrated Taxonomic Information System and uBio; and via the Catalogue of Life (see function \emph{col\_classification})

<<classification, eval=TRUE>>=
# As we already have Taxonomic Serial Numbers from ITIS, let's just get classifications from ITIS. Note that we could use uBio instead.
class_list <- classification(tsns)
sapply(class_list, nrow)

@@

\begin{figure}[!ht] 
\centering 
<<phylogeny, eval=TRUE, dev='png'>>=
phylogeny <- phylomatic_tree(taxa=as.character(allnames$spname), taxnames=TRUE, get = 'POST', informat='newick', method = "phylomatic",
    storedtree = "R20120829", taxaformat = "slashpath", outformat = "newick", clean = "true", parallel=TRUE)
phylogeny$tip.label <- capwords(phylogeny$tip.label, onlyfirst=TRUE)
plot(phylogeny)
@
\caption{a phylogeny...} 
\end{figure}

@@

\begin{figure}[!ht] 
\centering 
<<map, eval=TRUE, dev='png'>>=
library(rgbif)

@@

# Make a map
p <- rgbif::map(occurr_list)
p <- p + guides(col = guide_legend(nrow = 2, byrow = TRUE)) + theme(legend.position="bottom")
p
@
\caption{a map...} 
\end{figure}

\bibliography{apprefs}
\end{document}

>>>>

\documentclass[letterpaper,superscriptaddress,showkeys,longbibliography,10pt]{revtex4-1}
\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\hypersetup
{
colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

@@

<<resolvenames, eval=TRUE, cache=TRUE>>=
# The tnrs function accepts a vector of 1 or more
splist_tnrs <- tnrs(query=splist, getpost="POST", source_ = "iPlant_TNRS")

@@

<<commonnames, eval=TRUE, cache=TRUE>>=
tsns <- get_tsn(searchterm=splist, searchtype="sciname", verbose=FALSE)
comnames <- lapply(tsns, getcommonnamesfromtsn)

@@

Another common task is getting the taxonomic tree upstream from your study taxa. We often know what family or order our taxa are in, but it we often don't know the tribes, subclasses, and superfafamilies. taxize provides many avenues to getting classifications. Two of them are accesible via a single function (\emph{classification}): the Integrated Taxonomic Information System (ITIS) and National Center for Biotechnology Information (NCBI); and via the Catalogue of Life (see function \emph{col\_classification}):

<<classification, eval=TRUE, cache=TRUE>>=
# As we already have Taxonomic Serial Numbers from ITIS, let's just get classifications from ITIS. Note that we could use NCBI instead.
class_list <- classification(tsns)
sapply(class_list, nrow)

@@

\begin{figure}[!h] 
\centering 
<<phylogeny, eval=TRUE, dev='pdf', out.width='4in', out.height='4in', cache=TRUE, message=FALSE>>=
phylogeny <- phylomatic_tree(taxa=as.character(allnames$spname), taxnames=TRUE, get = 'POST', informat='newick', method = "phylomatic",
    storedtree = "R20120829", taxaformat = "slashpath", outformat = "newick", clean = "true", parallel=TRUE)
phylogeny$tip.label <- capwords(phylogeny$tip.label, onlyfirst=TRUE)
plot(phylogeny)
@
\caption{a phylogeny...} 
\end{figure}

@@

\begin{figure}[!ht] 
\centering 
<<map, eval=TRUE, dev='pdf', out.width='\\textwidth'>>=
library(rgbif)
library(ggplot2)

@@

# Make a map
p <- gbifmap(occurr_list) + 
  guides(col = guide_legend(title="", nrow = 3, byrow = TRUE)) + 
  theme(legend.position="bottom", legend.key = element_blank()) +
  coord_equal()
p
@
\caption{a map...} 
\end{figure}

\end{document}

### 2013-05-10 inst/doc/taxize_withcode_appA.tex
cached files, small changes to text, better graphics

\usepackage{alltt}
\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\usepackage{natbib}
\usepackage{footnote}
\hypersetup
{
colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}

@@

Another common task is getting the taxonomic tree upstream from your study taxa. We often know what family or order our taxa are in, but it we often don't know the tribes, subclasses, and superfafamilies. taxize provides many avenues to getting classifications. Two of them are accesible via a single function (\emph{classification}): the Integrated Taxonomic Information System and uBio; and via the Catalogue of Life (see function \emph{col\_classification})

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# As we already have Taxonomic Serial Numbers from ITIS, let's just get}
\hlcomment{# classifications from ITIS. Note that we could use uBio instead.}
class_list <- \hlfunctioncall{classification}(tsns)
\hlfunctioncall{sapply}(class_list, nrow)
\end{alltt}
\begin{verbatim}
 [1] 12 11 12 11 12 12 12 12 12 12
\end{verbatim}
\begin{alltt}

@@

\begin{figure}[!ht] 
\centering 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
phylogeny <- \hlfunctioncall{phylomatic_tree}(taxa = \hlfunctioncall{as.character}(allnames$spname), taxnames = TRUE, 
    get = \hlstring{"POST"}, informat = \hlstring{"newick"}, method = \hlstring{"phylomatic"}, storedtree = \hlstring{"R20120829"}, 
    taxaformat = \hlstring{"slashpath"}, outformat = \hlstring{"newick"}, clean = \hlstring{"true"}, parallel = TRUE)
\end{alltt}
\begin{verbatim}

@@

{\centering \includegraphics[width=\maxwidth]{figure/phylogeny}

@@

\hlcomment{# Make a map}
p <- \hlfunctioncall{gbifmap}(occurr_list) + \hlfunctioncall{guides}(col = \hlfunctioncall{guide_legend}(title = \hlstring{""}, nrow = 2, 
    byrow = TRUE)) + \hlfunctioncall{theme}(legend.position = \hlstring{"bottom"}, legend.key = \hlfunctioncall{element_blank}())
p
\end{alltt}
\end{kframe}

{\centering \includegraphics[width=\maxwidth]{figure/map}

@@

\bibliography{apprefs}
\end{document}

>>>>

\usepackage{alltt}
\usepackage[utf8]{inputenc}
\usepackage{color,dcolumn,graphicx,hyperref}
\hypersetup
{
colorlinks = true, linkcolor = blue, citecolor = blue, urlcolor = blue,
}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

@@

Another common task is getting the taxonomic tree upstream from your study taxa. We often know what family or order our taxa are in, but it we often don't know the tribes, subclasses, and superfafamilies. taxize provides many avenues to getting classifications. Two of them are accesible via a single function (\emph{classification}): the Integrated Taxonomic Information System (ITIS) and National Center for Biotechnology Information (NCBI); and via the Catalogue of Life (see function \emph{col\_classification}):

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcomment{# As we already have Taxonomic Serial Numbers from ITIS, let's just get}
\hlcomment{# classifications from ITIS. Note that we could use NCBI instead.}
class_list <- \hlfunctioncall{classification}(tsns)
\hlfunctioncall{sapply}(class_list, nrow)
\end{alltt}
\begin{verbatim}
 [1] 12 11 12 11 12 12 12 12 12 12
\end{verbatim}
\begin{alltt}

@@

\begin{figure}[!h] 
\centering 
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
phylogeny <- \hlfunctioncall{phylomatic_tree}(taxa = \hlfunctioncall{as.character}(allnames$spname), taxnames = TRUE, 
    get = \hlstring{"POST"}, informat = \hlstring{"newick"}, method = \hlstring{"phylomatic"}, storedtree = \hlstring{"R20120829"}, 
    taxaformat = \hlstring{"slashpath"}, outformat = \hlstring{"newick"}, clean = \hlstring{"true"}, parallel = TRUE)
\end{alltt}
\begin{verbatim}

@@

{\centering \includegraphics[width=4in,height=4in]{figure/phylogeny}

@@

\hlcomment{# Make a map}
p <- \hlfunctioncall{gbifmap}(occurr_list) + \hlfunctioncall{guides}(col = \hlfunctioncall{guide_legend}(title = \hlstring{""}, nrow = 3, 
    byrow = TRUE)) + \hlfunctioncall{theme}(legend.position = \hlstring{"bottom"}, legend.key = \hlfunctioncall{element_blank}()) + 
    \hlfunctioncall{coord_equal}()
p
\end{alltt}
\end{kframe}

{\centering \includegraphics[width=\textwidth]{figure/map}

@@

\end{document}
